Samsung Analysis
========================================================

Load Library
--------------------------------------------------------
```{r}
library(tree)
library(stringr)
library(randomForest)


```


Processing
--------------------------------------------------------
Download the data & read the data in 

```{r}
setwd("/Users/Carlos/Dropbox/Courses/Data Analysis/Exercices/Data Analysis/Data Analysis 2/code/SamsungAnalysis/")
getwd()
if (!file.exists("../../data/samsungData.rda")){
  download.file("https://spark-public.s3.amazonaws.com/dataanalysis/samsungData.rda", destfile="../../data/samsungData.rda", method = "curl")
  dateDownloaded <- date()
  dateDownloaded
}

load("../../data/samsungData.rda")
```
Munging Data
-------------------------------------------------------------

```{r}
punct <- c("\\(","\\)","\\-","\\,")
for (i in punct){ 
  names(samsungData) <- str_replace_all(names(samsungData),i, "_")
  }
samsungData$activity <- as.factor(samsungData$activity)
samsungData$activityNum <- as.numeric(samsungData$activity)
```
### Look at the data set
```{r}
#head(samsungData)
#summary(samsungData)
#sapply(samsungData[1, ], class)
table(samsungData$activity)
table(mapply(FUN=class, samsungData))
```

### Find out about missing values
```{r}
sum(is.na(samsungData))
```

Prediction
---------------------------------------------------------------
### 1. Find the right data


### 2. Define your error rate 

### 3. Split data into

#### Training
For this analysis the training set must include the data from subjects 1, 3, 5, and 6
```{r}
trainingSet <- samsungData[samsungData$subject == c(1, 3, 5, 6),]

```

#### Testing
The test set is the data from subjects 27, 28, 29, and 30
```{r}
testSet <- samsungData[samsungData$subject == c(27, 28, 29, 30),]

```


#### Validation (optional)
For this prediction, we don't use validation set of data

### 4. On the training set pick features


__Prediction Tree for Laying__ 

```{r fig.width=12, fig.height=6}

activity <- "laying"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]

#layingTree <- tree(activity ~ 
#                tGravityAcc_mean___X
            #  + tGravityAcc_mean___Y
            #    + tBodyAcc_arCoeff___Y_1
            #  + tBodyAcc_correlation___Y_Z
#             ,
#              data=trainingSet)

layingTree <- randomForest(activity ~ 
                             tGravityAcc_mean___X,
             data=trainingSet, prox=TRUE, ntree=50)


tree1 <- layingTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)

```
__Prediction Tree for Sitting__ 

```{r fig.width=12, fig.height=6}

activity <- "sitting"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]

sittingTree <- randomForest(activity ~ 
                tGravityAcc_mean___Y
            #  + tGravityAcc_min___Y
            #   +tGravityAcc_max___Y
             #   + tBodyAcc_correlation___X_Y
              ,
              data=trainingSet, prox=TRUE, ntree=50)


tree1 <- sittingTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)


```
__Prediction Tree for Standing__ 

```{r fig.width=12, fig.height=6}

activity <- "standing"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]
standingTree <- randomForest(activity ~ 
             angle_Y_gravityMean_,
              data=trainingSet, prox=TRUE, ntree=50)
tree1 <- sittingTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
#text(tree1)
set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))
print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)
```
__Prediction Tree for Walk__ 

```{r fig.width=12, fig.height=6}

activity <- "walk"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]

par(mfrow=c(1,3))
boxplot(trainingSet$tBodyAccJerk_energy___X ~ trainingSet$activity)
boxplot(trainingSet$fBodyAcc_max___X ~ trainingSet$activity)
boxplot(trainingSet$fBodyBodyAccJerkMag_mean__ ~ trainingSet$activity)

walkTree <- randomForest(activity ~ 
                               tBodyAccJerk_energy___X
                               
                               + fBodyAcc_max___X
                               #fBodyAccJerk_bandsEnergy___1_8
                             #+
                              #fBodyAccJerk_maxInds_X
          #   fBodyBodyAccJerkMag_mean__
           #                  +fBodyAccJerk_bandsEnergy___1_16
              ,
              data=trainingSet, prox=TRUE, ntree=50)
tree1 <- walkTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
#text(tree1)
set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)
```

__Prediction Tree for Walkdown__ 


```{r fig.width=12, fig.height=6}
activity <- "walkdown"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]
par(mfrow=c(1,3))
boxplot(trainingSet$tBodyAcc_energy___X ~ trainingSet$activity)
boxplot(trainingSet$tBodyAcc_std___X ~ trainingSet$activity)
boxplot(trainingSet$tBodyAccJerk_max___X ~ trainingSet$activity)

walkdownTree <- randomForest(activity ~ 
                               tBodyAccJerk_max___X
                              + tBodyAcc_energy___X
                             #+  tBodyAcc_std___X
                            #+ tBodyAcc_mad___X
                            # + tBodyAccJerk_max___X
              ,
              data=trainingSet, prox=TRUE, ntree=50)
tree1 <- walkdownTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
#text(tree1)

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)
```
__Prediction Tree for Walkup__

```{r fig.width=12, fig.height=6}
activity <- "walkup"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]
walkup <- randomForest(activity ~ 
                         tGravityAcc_energy___Y      
                       #  tGravityAcc_mean___Y
                       + tBodyAccJerk_std___X
                       + tBodyAccJerk_std___Z
                      #+ tGravityAcc_max___Y
              ,
              data=trainingSet, prox=TRUE, ntree=50)
tree1 <- walkup
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
#text(tree1)

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)
```


### 5. On the training set pick prediction function

```{r}
combineTree <- randomForest(activity ~ 
                tGravityAcc_mean___Y # laying
                + tGravityAcc_mean___X # sitting
                + angle_Y_gravityMean_  # standing
                + tBodyAccJerk_energy___X #walk
                + fBodyBodyAccJerkMag_mean__ #walk*
                + tBodyAccJerk_max___X #walkdown
                + tBodyAcc_energy___X #walkdown 
                + tBodyAccJerk_std___X # walkup
                + tBodyAccJerk_std___Z # walkup
              ,
              data=trainingSet, prox=TRUE, ntree=50)

tree1 <- combineTree


summary(tree1)
par(mfrow=c(1,1))
plot(tree1, main="Error rates of trees") 
```



### 6. On the training set cross-validate


```{r}




```


### 7. If no validation - apply 1x to test set

```{r}
set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))


errorRates <- data.frame(stringsAsFactors=TRUE)
activities <- levels(testSet$activity)

for(a in activities){
#  print(c("Sensitivity:",a,":",sum(a == predict(tree1, testSamples[testSamples$activity==a,], type="class"))/sum(testSamples$activity==a)))
  
  sensitivity <- sum(a == predict(tree1, testSamples[testSamples$activity==a,], type="class"))/sum(testSamples$activity==a)
  
  specificity <- sum(a != predict(tree1, testSamples[testSamples$activity!=a,], type="class"))/sum(testSamples$activity!=a)
  
  falsePositive <- sum( a == predict(tree1, testSamples[testSamples$activity!=a,], type="class"))/length(testSamples$activity)
  
  falseNegative <- sum(a != predict(tree1, testSamples[testSamples$activity==a,], type="class"))/length(testSamples$activity)
  
  errorRates <- rbind(errorRates, c( factor(a, levels=levels(testSet$activity)) , sensitivity, specificity, falsePositive, falseNegative))
  
#  print(c("Specificity:", a, ":", sum(activity != predict(tree1, testSamples[testSamples$activity!=a,], type="class"))/sum(testSamples$activity!=a) ))
  
}
colnames(errorRates)<-c("activity","sensitivity","specificity","falsePositive", "falseNegative")

errorRates

```




### 8. If validation - apply to test set and refine 

### 9. If validation - apply 1x to validation
