Samsung Analysis
========================================================

Load Library
--------------------------------------------------------


Processing
--------------------------------------------------------
Download the data & read the data in 

```{r}
setwd("/Users/Carlos/Dropbox/Courses/Data Analysis/Exercices/Data Analysis/Data Analysis 2/code/SamsungAnalysis/")
getwd()
if (!file.exists("../../data/samsungData.rda")){
  download.file("https://spark-public.s3.amazonaws.com/dataanalysis/samsungData.rda", destfile="../../data/samsungData.rda", method = "curl")
  dateDownloaded <- date()
  dateDownloaded
}

load("../../data/samsungData.rda")
```
Munging Data
-------------------------------------------------------------
```{r}
samsungData$activity <- as.factor(samsungData$activity)

```
### Look at the data set
```{r}
#head(samsungData)
summary(samsungData)
sapply(samsungData[1, ], class)
table(samsungData$activity)
```

### Find out about missing values
```{r}
sum(is.na(samsungData))
```

Prediction
---------------------------------------------------------------
### 1. Find the right data


### 2. Define your error rate 

### 3. Split data into

#### Training
For this analysis the training set must include the data from subjects 1, 3, 5, and 6
```{r}
trainingSet <- samsungData[samsungData$subject == c(1, 3, 5, 6),]

```

#### Testing
The test set is the data from subjects 27, 28, 29, and 30
```{r}
testSet <- samsungData[samsungData$subject == c(27, 28, 29, 30),]

```


#### Validation (optional)
For this prediction, we don't use validation set of data

### 4. On the training set pick features

#### Singular value decomposition
```{r}
numericActivity <- as.numeric(as.factor(trainingSet$activity))
# Singular value decomposition
svd1 = svd(scale(trainingSet[ ,-c(562,563)])) 
par(mfrow=c(1,2))
plot(svd1$u[,1],col=numericActivity,pch=19) 
plot(svd1$u[,2],col=numericActivity,pch=19)
```


### 5. On the training set pick prediction function

### 6. On the training set cross-validate

### 7. If no validation - apply 1x to test set

### 8. If validation - apply to test set and refine 

### 9. If validation - apply 1x to validation
