Samsung Analysis
========================================================

Load Library
--------------------------------------------------------
```{r}
library(tree)
library(stringr)
library(randomForest)


```


Processing
--------------------------------------------------------
Download the data & read the data in 

```{r}
setwd("/Users/Carlos/Dropbox/Courses/Data Analysis/Exercices/Data Analysis/Data Analysis 2/code/SamsungAnalysis/")
getwd()
if (!file.exists("../../data/samsungData.rda")){
  download.file("https://spark-public.s3.amazonaws.com/dataanalysis/samsungData.rda", destfile="../../data/samsungData.rda", method = "curl")
  dateDownloaded <- date()
  dateDownloaded
}

load("../../data/samsungData.rda")
```
Munging Data
-------------------------------------------------------------

```{r}
punct <- c("\\(","\\)","\\-","\\,")
for (i in punct){ 
  names(samsungData) <- str_replace_all(names(samsungData),i, "_")
  }
samsungData$activity <- as.factor(samsungData$activity)
samsungData$activityNum <- as.numeric(samsungData$activity)
```
### Look at the data set
```{r}
#head(samsungData)
#summary(samsungData)
#sapply(samsungData[1, ], class)
table(samsungData$activity)
table(mapply(FUN=class, samsungData))
```

### Find out about missing values
```{r}
sum(is.na(samsungData))
```

Prediction
---------------------------------------------------------------
### 1. Find the right data


### 2. Define your error rate 

### 3. Split data into

#### Training
For this analysis the training set must include the data from subjects 1, 3, 5, and 6
```{r}
trainingSet <- samsungData[samsungData$subject == c(1, 3, 5, 6),]

```

#### Testing
The test set is the data from subjects 27, 28, 29, and 30
```{r}
testSet <- samsungData[samsungData$subject == c(27, 28, 29, 30),]

```


#### Validation (optional)
For this prediction, we don't use validation set of data

### 4. On the training set pick features

#### Singular value decomposition:
```{r fig.width=12, fig.height=6}
#numericActivity <- as.numeric(as.factor(trainingSet$activity))
# Singular value decomposition
svd1 = svd(scale(trainingSet[ ,-c(562,563,564)])) 
par(mfrow=c(1,2))
plot(svd1$u[,1],col=trainingSet$activityNum,pch=19) 
plot(svd1$u[,2],col=trainingSet$activityNum,pch=19)
```

Find maximum contributor:
```{r}
# Find maximum contributor
par(mfrow=c(1,1))
plot(svd1$v[,2],pch=19)
```

```{r}
# Find maximum contributor
par(mfrow=c(1,1))
plot(svd1$d^2/sum(svd1$d^2),pch=19)
```


New clustering with maximum contributer:
```{r}
source("http://dl.dropbox.com/u/7710864/courseraPublic/myplclust.R")
# New clustering with maximum contributer
maxContrib <- which.max(svd1$v[,2])
distanceMatrix <- dist(trainingSet[ ,c(10:12,maxContrib)]) 
hclustering <- hclust(distanceMatrix) 
myplclust(hclustering,lab.col=numericActivity)
names(trainingSet)[maxContrib]
```
#### Pick features on the training set:
10 names of variables ordered by contribution to prediction

```{r}
maxContrib <- which.max(svd1$v[,2])
features<-order(abs(svd1$v[,2]), decreasing=TRUE)
head(features, n=20)
```

#### Prediction Trees

__Prediction Tree 1.__ 
* outcome: activity
* covariantes: 
  * fBodyAcc_meanFreq___Z
  
```{r fig.width=12, fig.height=6}
par(mfrow=c(1,1))
plot(trainingSet$activityNum, trainingSet[,296], pch=19)

tree1 <- tree(activity ~ trainingSet[,296], 
              data=trainingSet)
summary(tree1)
plot(tree1) 
text(tree1)
table(testSamples$activity,predict(tree1, testSamples, type="class"))

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]

sum(testSamples$activity == predict(tree1, testSamples, type="class"))/length(testSamples)


```
__Prediction Tree 2__ 
* outcome: activity
* covariantes: 
  * fBodyAcc_meanFreq___Z
  * tBodyGyroMag_arCoeff__1

```{r fig.width=12, fig.height=6}

plot(trainingSet[,296],trainingSet[,249], col=trainingSet$activityNum, pch=19)

tree1 <- tree(activity ~ trainingSet[,296] +  trainingSet[,249], 
              data=trainingSet)
summary(tree1)
plot(tree1) 
text(tree1)
table(testSamples$activity,predict(tree1, testSamples, type="class"))

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
sum(testSamples$activity == predict(tree1, testSamples, type="class"))/length(testSamples)
```

```{r fig.width=12, fig.height=6}

tree1 <- tree(activity ~ trainingSet[,296] +  trainingSet[,249] + trainingSet[,34] + trainingSet[,250], 
              data=trainingSet)
summary(tree1)
plot(tree1) 
text(tree1)
table(testSamples$activity,predict(tree1, testSamples, type="class"))

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
sum(testSamples$activity == predict(tree1, testSamples, type="class"))/length(testSamples)
```


```{r fig.width=12, fig.height=6}

tree1 <- tree(activity ~ trainingSet[,296] +  trainingSet[,249] + trainingSet[,34] + trainingSet[,250] + trainingSet[,375] + trainingSet[,76] + trainingSet[,75] + trainingSet[,77] + trainingSet[,74] + trainingSet[,513] + trainingSet[,210] + trainingSet[,223] + trainingSet[,453] + trainingSet[,211] + trainingSet[,224] + trainingSet[,72] + trainingSet[,114] + trainingSet[,73] + trainingSet[,71] + trainingSet[,295], 
              data=trainingSet)
summary(tree1)
plot(tree1) 
text(tree1)
table(testSamples$activity,predict(tree1, testSamples, type="class"))

pruneTree <- prune.tree(tree1, best=7)
plot(pruneTree) 
text(pruneTree)

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
sum(testSamples$activity == predict(pruneTree, testSamples, type="class"))/length(testSamples)

table(testSamples$activity,predict(pruneTree, testSamples, type="class"))

```



```{r fig.width=12, fig.height=6}

tree1 <- randomForest(activity ~ trainingSet[,296] +  trainingSet[,249] + trainingSet[,34] + trainingSet[,250] + trainingSet[,375] + trainingSet[,76] + trainingSet[,75] + trainingSet[,77] + trainingSet[,74] + trainingSet[,513] + trainingSet[,210] + trainingSet[,223] + trainingSet[,453] + trainingSet[,211] + trainingSet[,224] + trainingSet[,72] + trainingSet[,114] + trainingSet[,73] + trainingSet[,71] + trainingSet[,295], 
              data=trainingSet, prox=TRUE, ntree=500)
summary(tree1)
plot(tree1) 
text(tree1)

testSamples <- testSet[sample(1:371,size=328,replace=T),]
getTree(tree1)
table(testSamples$activity,predict(tree1, testSamples))

sum(testSamples$activity == predict(tree1, testSamples))/length(testSamples)

pruneTree <- prune.tree(tree1, best=7)
plot(pruneTree) 
text(pruneTree)

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
sum(testSamples$activity == predict(pruneTree, testSamples, type="class"))/length(testSamples)

table(testSamples$activity,predict(pruneTree, testSamples, type="class"))

```

__Prediction Tree 3__ 
* outcome: activity
* covariantes: 
  * fBodyAcc_meanFreq___Z
  * tBodyGyroMag_arCoeff__1
  * tBodyAcc_arCoeff___Z_1

```{r fig.width=12, fig.height=6}

tree1 <- tree(activity ~ fBodyAcc_meanFreq___Z + tBodyGyroMag_arCoeff__1 + tBodyAcc_arCoeff___Z_1, 
              data=trainingSet)
summary(tree1)
plot(tree1) 
text(tree1)

sum(testSet$activity == predict(tree1, testSet, type="class"))/length(trainingSet)
```

__Prediction Tree 4__ 
* outcome: activity
* covariantes: 
  * fBodyAcc_meanFreq___Z
  * tBodyGyroMag_arCoeff__1
  * tBodyAcc_arCoeff___Z_1
  * tBodyGyroMag_arCoeff__2

```{r fig.width=12, fig.height=6}

tree1 <- tree(activity ~ fBodyAcc_meanFreq___Z + tBodyGyroMag_arCoeff__1 + tBodyAcc_arCoeff___Z_1 + tBodyGyroMag_arCoeff__2, 
              data=trainingSet)
summary(tree1)
plot(tree1) 
text(tree1)

sum(testSet$activity == predict(tree1, testSet, type="class"))/length(trainingSet)
```


__Prediction Tree 4__ 
* outcome: activity
* covariantes: 
  * all variables

```{r fig.width=12, fig.height=6}

tree1 <- tree(activity ~ ., 
              data=trainingSet)
summary(tree1)
plot(tree1) 
text(tree1)

sum(testSet$activity == predict(tree1, testSet, type="class"))/length(trainingSet)
```
### Poison Regression

```{r}
glm1 <- glm(activityNum ~ fBodyAcc_meanFreq___Z + tBodyGyroMag_arCoeff__1 + tBodyAcc_arCoeff___Z_1 + tBodyGyroMag_arCoeff__2, data = trainingSet)

sum(testSet$activityNum == floor(predict(glm1, testSet)))/length(trainingSet)


lm1 <- lm(activityNum ~ fBodyAcc_meanFreq___Z + tBodyGyroMag_arCoeff__1 + tBodyAcc_arCoeff___Z_1 + tBodyGyroMag_arCoeff__2, data = trainingSet)

sum(testSet$activityNum == floor(predict(lm1, testSet)))/length(trainingSet)


```





#### K-means clustering
```{r}
kClust <- kmeans(trainingSet[ ,-c(562,563)],centers=6,nstart=100) 
table(kClust$cluster,trainingSet$activity)
```

Custer 1 Variable Centers(Laying)
```{r}
par(mfrow=c(1,1))
plot(kClust$center[1,1:10],pch=19,ylab="Cluster Center",xlab="")
```

Custer 2 Variable Centers(Walkin)
```{r}
par(mfrow=c(1,1))
plot(kClust$center[6,1:10],pch=19,ylab="Cluster Center",xlab="")
```




### 5. On the training set pick prediction function

### 6. On the training set cross-validate

```{r}

```


### 7. If no validation - apply 1x to test set

### 8. If validation - apply to test set and refine 

### 9. If validation - apply 1x to validation


```{r}
library(splines)
ns1 <- ns(trainingSet$fBodyAcc_meanFreq___Z,df=1)
par(mfrow=c(1,3))
plot(trainingSet$activityNum,ns1[,1])
plot(trainingSet$activityNum,ns1[,2])
plot(trainingSet$activityNum,ns1[,3])

lm1 <- lm(trainingSet$activityNum ~ ns1)
summary(lm1)

sum(testSamples$activityNum == round(predict(lm1, testSamples)))/length(testSamples)

table(testSamples$activityNum,round(predict(lm1, testSamples)))

```

```{r}



```

```{r fig.width=12, fig.height=6}

# Singular value decomposition


tree1 <- tree(activity ~ trainingSet$angle_Y_gravityMean_ + trainingSet$fBodyAccJerk_bandsEnergy___1_8 + trainingSet$tGravityAcc_min___Y, 
              data=trainingSet)
summary(tree1)
plot(tree1) 
text(tree1)



set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))
sum(testSamples$activity == predict(tree1, testSamples, type="class"))/length(testSamples)

plot(colMeans(trainingSet[trainingSet$activityNum==6,-c(562,563,564)]), pch=19)

```

### Laying

```{r fig.width=12, fig.height=6}

activity <- "laying"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]

#layingTree <- tree(activity ~ 
#                tGravityAcc_mean___X
            #  + tGravityAcc_mean___Y
            #    + tBodyAcc_arCoeff___Y_1
            #  + tBodyAcc_correlation___Y_Z
#             ,
#              data=trainingSet)

layingTree <- randomForest(activity ~ 
                             tGravityAcc_mean___X,
             data=trainingSet, prox=TRUE, ntree=50)


tree1 <- layingTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
text(tree1)



set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)


```


### Sitting

```{r fig.width=12, fig.height=6}

activity <- "sitting"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]

sittingTree <- randomForest(activity ~ 
                tGravityAcc_mean___Y
            #  + tGravityAcc_min___Y
            #   +tGravityAcc_max___Y
             #   + tBodyAcc_correlation___X_Y
              ,
              data=trainingSet, prox=TRUE, ntree=50)


tree1 <- sittingTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
text(tree1)



set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)


```

### Standing

```{r fig.width=12, fig.height=6}

activity <- "standing"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]



standingTree <- randomForest(activity ~ 
             angle_Y_gravityMean_
              ,
              data=trainingSet, prox=TRUE, ntree=50)


tree1 <- sittingTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
#text(tree1)



set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)


```
### Walk

```{r fig.width=12, fig.height=6}

activity <- "walk"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]



walkTree <- randomForest(activity ~ 
                               tBodyAccJerk_energy___X
                               
                               + fBodyAcc_max___X
                               #fBodyAccJerk_bandsEnergy___1_8
                             #+
                              #fBodyAccJerk_maxInds_X
          #   fBodyBodyAccJerkMag_mean__
           #                  +fBodyAccJerk_bandsEnergy___1_16
              ,
              data=trainingSet, prox=TRUE, ntree=50)


tree1 <- walkTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
#text(tree1)



set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)
```


### Walkdown

```{r fig.width=12, fig.height=6}
activity <- "walkdown"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]
walkdownTree <- randomForest(activity ~ 
                               tBodyAcc_energy___X
                             +  tBodyAcc_std___X
                            + tBodyAcc_mad___X
                            # + tBodyAccJerk_max___X
              ,
              data=trainingSet, prox=TRUE, ntree=50)
tree1 <- walkdownTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
#text(tree1)

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)
```

### Walkup

```{r fig.width=12, fig.height=6}
activity <- "walkup"
trainingSetActivity <- trainingSet[ trainingSet$activity==activity,]
walkup <- randomForest(activity ~ 
                         tGravityAcc_energy___Y      
                       #  tGravityAcc_mean___Y
                       + tBodyAccJerk_std___X
                       + tBodyAccJerk_std___Z
                      #+ tGravityAcc_max___Y
              ,
              data=trainingSet, prox=TRUE, ntree=50)
tree1 <- walkup
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
#text(tree1)

set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

print(c("Sensitivity:",activity))
sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)
print(c("Specificity:", activity))
sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity)
```


### Combining randomForest
```{r}
combineTree <- randomForest(activity ~ 
                tGravityAcc_mean___Y # laying
                + tGravityAcc_mean___X # sitting
                + angle_Y_gravityMean_  # standing
                + tBodyAccJerk_energy___X #walk
                + fBodyAcc_max___X #walk
                + tBodyAcc_energy___X #walkdown
                + tBodyAcc_std___X #walkdown
                + tBodyAcc_mad___X #walkdown
               # + tGravityAcc_energy___Y     # walkup 
                + tBodyAccJerk_std___X # walkup
                + tBodyAccJerk_std___Z # walkup
              ,
              data=trainingSet, prox=TRUE, ntree=50)

tree1 <- combineTree
summary(tree1)
par(mfrow=c(1,1))
plot(tree1) 
#text(tree1)
set.seed(333)
testSamples <- testSet[sample(1:371,size=328,replace=T),]
table(testSamples$activity,predict(tree1, testSamples, type="class"))

for(activity in c("laying","sitting", "standing","walk","walkdown","walkup")){
  print(c("Sensitivity:",activity,":",sum(activity == predict(tree1, testSamples[testSamples$activity==activity,], type="class"))/sum(testSamples$activity==activity)))
  print(c("Specificity:", activity, ":", sum(activity != predict(tree1, testSamples[testSamples$activity!=activity,], type="class"))/sum(testSamples$activity!=activity) ))
}


```

