Data Analysis - Week 6
========================================================

Prediction Study Design
--------------------------------------------------------
### Steps in building a prediction
1. Find the right data
2. Define your error rate 
3. Split data into:
  * Training
  * Testing
  * Validation (optional)
4. On the training set pick features
5. On the training set pick prediction function 
6. On the training set cross-validate
7. If no validation - apply 1x to test set
8. If validation - apply to test set and refine 
9. If validation - apply 1x to validation



Cross Validation
--------------------------------------------------------
### Overfitting
```{r}
set.seed(12345)
x <- rnorm(10); y <- rnorm(10); z <- rbinom(10,size=1,prob=0.5) 
plot(x,y,pch=19,col=(z+3))
```

### Classifier
Example predictor function: If -0.2 < y < 0.6 call blue, otherwise green

```{r fig.width=10, fig.height=6}

par(mfrow=c(1,2))
zhat<-(-0.2<y)&(y<0.6)
plot(x,y,pch=19,col=(z+3)) 
plot(x,y,pch=19,col=(zhat+3))

```

### New Data
Example predictor function: If -0.2 < y < 0.6 call blue, otherwise green
It doesn't work with new data
```{r fig.width=10, fig.height=6}
set.seed(1233)
xnew <- rnorm(10); ynew <- rnorm(10); znew <- rbinom(10,size=1,prob=0.5) 
par(mfrow=c(1,2)); zhatnew <- (-0.2 < ynew) & (ynew < 0.6) 
plot(xnew,ynew,pch=19,col=(z+3)); plot(xnew,ynew,pch=19,col=(zhatnew+3))
```

### Subsampling example

```{r fig.width=10, fig.height=6}
y1 <- y[1:5]; x1 <- x[1:5]; z1 <- z[1:5]
y2 <- y[6:10]; x2 <- x[6:10]; z2 <- z[6:10];
zhat2<-(y2<1)&(y2>-0.5)
par(mfrow=c(1,3))
plot(x1,y1,col=(z1+3),pch=19); plot(x2,y2,col=(z2+3),pch=19); plot(x2,y2,col=(zhat2+3),pch=19)

```

Predicting with regression models
--------------------------------------------------------------

### Example: Old faithful eruptions
```{r}
data(faithful) 
dim(faithful)
set.seed(333)
trainSamples <- sample(1:272,size=(272/2),replace=F) 
trainFaith <- faithful[trainSamples,]
testFaith <- faithful[-trainSamples,] 
head(trainFaith)

```

#### Eruption duration versus waiting time
```{r}
plot(trainFaith$waiting,trainFaith$eruptions,pch=19,col="blue",xlab="Waiting",ylab="Duration")

```
#### Fit linear model
ED = b0 + b1 WT + e
```{r}
lm1 <- lm(eruptions ~ waiting,data=trainFaith)
summary(lm1)
```
#### Model Fit
```{r}
plot(trainFaith$waiting,trainFaith$eruptions,pch=19,col="blue",xlab="Waiting",ylab="Duration") 
lines(trainFaith$waiting,lm1$fitted,lwd=3)

```

#### Predict New Value

```{r}
coef(lm1)[1] + coef(lm1)[2]*80

newdata <- data.frame(waiting=80) 
predict(lm1,newdata)
```

#### Plot predictions - training and test

```{r fig.width=10, fig.height=6}
par(mfrow=c(1,2)) 
plot(trainFaith$waiting,trainFaith$eruptions,pch=19,col="blue",xlab="Waiting",ylab="Duration") 
lines(trainFaith$waiting,predict(lm1),lwd=3) 
plot(testFaith$waiting,testFaith$eruptions,pch=19,col="blue",xlab="Waiting",ylab="Duration") 
lines(testFaith$waiting,predict(lm1,newdata=testFaith),lwd=3)

```

### Get training set/test set errors
```{r}
# Calculate RMSE on training
sqrt(sum((lm1$fitted-trainFaith$eruptions)^2))

# Calculate RMSE on test
sqrt(sum((predict(lm1,newdata=testFaith)-testFaith$eruptions)^2))

```

#### Prediction intervals
```{r}
pred1 <- predict(lm1,newdata=testFaith,interval="prediction")
ord <- order(testFaith$waiting) 
plot(testFaith$waiting,testFaith$eruptions,pch=19,col="blue") 
matlines(testFaith$waiting[ord],pred1[ord,],type="l",,col=c(1,2,2),lty = c(1,1,1), lwd=3)

```

### Example with binary data: Baltimore Ravens

#### Ravens Data
```{r}
##download.file("https://dl.dropbox.com/u/7710864/data/ravensData.rda", destfile="./data/ravensData.rda",method="curl")
load("./data/ravensData.rda") 
head(ravensData)
```
#### Fit a logistic regresion
logit(E[RWi|RSi]) = b0 + b1RSi

```{r fig.width=10, fig.height=6}
glm1 <- glm(ravenWinNum ~ ravenScore,family="binomial",data=ravensData) 
par(mfrow=c(1,2))
boxplot(predict(glm1) ~ ravensData$ravenWinNum,col="blue")
boxplot(predict(glm1,type="response") ~ ravensData$ravenWinNum,col="blue")
```

#### Choosing a cutoff (re-substitution)
```{r}
par(mfrow=c(1,1))
xx <- seq(0,1,length=10); err <- rep(NA,10) 
for(i in 1:length(xx)){
  err[i] <- sum((predict(glm1,type="response") > xx[i]) != ravensData$ravenWinNum) 
  }
plot(xx,err,pch=19,xlab="Cutoff",ylab="Error")
```

#### Comparing models with cross validation
```{r}
library(boot)
cost <- function(win, pred = 0) mean(abs(win-pred) > 0.5)
glm1 <- glm(ravenWinNum ~ ravenScore,family="binomial",data=ravensData) 
glm2 <- glm(ravenWinNum ~ ravenScore,family="gaussian",data=ravensData) 
cv1 <- cv.glm(ravensData,glm1,cost,K=3)
cv2 <- cv.glm(ravensData,glm2,cost,K=3)
cv1$delta
cv2$delta
```

Predicting with trees
---------------------------------------------------------------
### Key Ideas
* Iteratively split variables into groups
* Split where maximally predictive
* Evaluate "homogeneity" within each branch
* Fitting multiple trees often works better (forests)
Pros:
* Easy to implement
* Easy to interpret
* Better performance in nonlinear settings
Cons:
* Without pruning/cross-validation can lead to overfitting
* Harder to estimate uncertainty
* Results may be variable

### Basic algorithm
1. Start with all variables in one group
2. Find the variable/split that best separates the outcomes
3. Divide the data into two groups ("leaves") on that split ("node")
4. Within each split, find the best variable/split that separates the outcomes 
5. Continue until the groups are too small or sufficiently "pure"

### Example: Iris Data
```{r}
data(iris) 
names(iris)
table(iris$Species)

```

#### Iris petal widths/sepal width
```{r}
par(mfcol=c(1,1))
plot(iris$Petal.Width,iris$Sepal.Width,pch=19,col=as.numeric(iris$Species)) 
legend(1,4.5,legend=unique(iris$Species),col=unique(as.numeric(iris$Species)),pch=19)
```

#### Iris petal widths/sepal width
```{r}
# An alternative is library(rpart)
library(tree)
tree1 <- tree(Species ~ Sepal.Width + Petal.Width,data=iris) 
summary(tree1)
```

#### Plot tree
```{r}
plot(tree1) 
text(tree1)
```

#### Another way of looking at a CART model
```{r}
plot(iris$Petal.Width,iris$Sepal.Width,pch=19,col=as.numeric(iris$Species)) 
partition.tree(tree1,label="Species",add=TRUE) 
legend(1.75,4.5,legend=unique(iris$Species),col=unique(as.numeric(iris$Species)),pch=19)

```

#### Predicting new values
```{r}
set.seed(32313)
newdata <- data.frame(Petal.Width = runif(20,0,2.5),Sepal.Width = runif(20,2,4.5)) 
pred1 <- predict(tree1,newdata)
pred1
```

#### Overlaying new values
```{r}
pred1 <- predict(tree1,newdata,type="class") 
plot(newdata$Petal.Width,newdata$Sepal.Width,col=as.numeric(pred1),pch=19) 
partition.tree(tree1,"Species",add=TRUE)

```

### Pruning trees example: Cars
```{r}
data(Cars93,package="MASS") 
head(Cars93)

```
#### Build a tree
```{r}
treeCars <- tree(DriveTrain ~ MPG.city + MPG.highway + AirBags + EngineSize + Width + Length + Weight + Price + Cylinders + Horsepower + Wheelbase,data=Cars93)
plot(treeCars) 
text(treeCars)
```
#### Plot errors
```{r fig.width=10, fig.height=6}
par(mfrow=c(1,2)) 
plot(cv.tree(treeCars,FUN=prune.tree,method="misclass")) 
plot(cv.tree(treeCars))

```

#### Prune the tree
```{r fig.width=10, fig.height=6}
par(mfrow=c(1,1)) 
pruneTree <- prune.tree(treeCars,best=4) 
plot(pruneTree)
text(pruneTree)
```

#### Show resubstitution error ∗
```{r}
table(Cars93$DriveTrain,predict(pruneTree,type="class"))
table(Cars93$DriveTrain,predict(treeCars,type="class"))
```
