Weekly Quiz 8
========================================================

### Question 1

Suppose this is the result of 85 hypothesis tests: 
What is the (observed) rate of false discoveries? What is the (observed) rate of false positives?

1. False discovery rate = 0.17 False positive rate = 0.33
2. False discovery rate = 0.33 False positive rate = 0.17
3. False discovery rate = 0.25 False positive rate = 0.10
4. False discovery rate = 0.20 False positive rate = 0.09 ✔




✘✔

```{r}

```
### Question 2
Generate P-values according to the following code:
```
set.seed(3343)
pValues = rep(NA,100)
for(i in 1:100){
  z = rnorm(20)
  x = rnorm(20)
  y = rnorm(20,mean=0.5*x)
  pValues[i] = summary(lm(y ~ x))$coef[2,4]
}
```
How many are significant at the alpha = 0.1 level when controlling the family wise error rate using the methods described in the lectures? When controlling the false discovery rate at the alpha = 0.1 level as described in the lectures?

1. FWER = 7 FDR = 61
2. FWER = 61 FDR = 7
3. FWER = 32 FDR = 5
5. FWER = 3 FDR = 5


```{r}

set.seed(3343)
pValues = rep(NA,100)
for(i in 1:100){
  z = rnorm(20)
  x = rnorm(20)
  y = rnorm(20,mean=0.5*x)
  pValues[i] = summary(lm(y ~ x))$coef[2,4]
  }

sum(pValues < 0.1/100)

  
  


```      
       

### Question 3
Suppose I want to generate data from the following model with a simulation:

```
y = b0 + b1*x + b2*z + e
```
where b0=1, b1=2, b2=3 and x, z, and e are normally distributed. Which one of the following is not a step in the simulation process?

1. Generate the y-values by adding yfit+e
2. Generate x, z, and e using rnorm()
3. Generate a random sample of values for b_0, b_1, and b_2
4. Generate the fitted values by adding yfit = 1 + 2*x + 3*z

✘✔
```{r}

```

### Question 4
Suppose data are generated from a model:

```
y = b0 + b1*x + e
```
where b0=1, b1=2 and x and e both have a normal distribution with mean zero and variance one. After the data are created, some data are lost. Use the lm() function in base R for model fitting. Case 1: Build a simulation where all values of y are observed but higher values of x are likely to be missing. Does the estimate of b1 change on average? If so how? Case 2 Build a simulation where all values of x are observed but higher values of y are likely to be missing. Does the estimate of b1 change on average? If so how?

1. Case 1: b1 is overestimated Case 2: b1 is overestimated
2. Case 1: b1 is estimated correctly Case 2: b1 is underestimated
3. Case 1: b1 is underestimated Case 2: b1 is underestimated
4. Case 1: b1 is estimated correctly Case 2: b1 is estimated correctly

```{r}

```


### Question 5
Exactly as in the last question, suppose data are generated from a model: y = b0 + b1*x + e where b0=1, b1=2 and x and e both have a normal distribution with mean zero and variance one. After the data are created, some data are lost. Answer the same questions below, but this time, use the rlm() function in the MASS package to fit the linear model instead of the lm() function in base R. Case 1: Build a simulation where all values of y are observed but higher values of x are likely to be missing. Does the estimate of b1 change on average? If so how? Case 2 Build a simulation where all values of x are observed but higher values of y are likely to be missing. Does the estimate of b1 change on average? If so how?

1. Case 1: b1 is underestimated Case 2: b1 is underestimated
2. Case 1: b1 is estimated correctly Case 2: b1 is underestimated
3. Case 1: b1 is overestimated Case 2: b1 is overestimated
4. Case 1: b1 is estimated correctly Case 2: b1 is estimated correctly
```{r}

```



