Raw Interest Rate Analysis
========================================================

Load Library
--------------------------------------------------------


Processing
--------------------------------------------------------
Download the data, read the data in and save as an RDA file

```{r}
getwd()
#download.file("https://spark-public.s3.amazonaws.com/dataanalysis/loansData.csv", destfile="../../data/loansData.csv", method = "curl")
dateDownloaded <- date()
dateDownloaded
loans <- read.csv("../../data/loansData.csv",stringsAsFactors=FALSE)
```
Munging Data
-------------------------------------------------------------
```{r}
#Convert Interest Rate to numeric
loans$Interest.Rate<-as.numeric(gsub("%","",loans$Interest.Rate))
#Convert Loan Lenght to numeric
loans$Loan.Length <- as.numeric(gsub("months","",loans$Loan.Length))
#Factor Loan Purpose
loans$Loan.Purpose <- as.factor(loans$Loan.Purpose)
#Convert Debt to Income Ratio to numeric
loans$Debt.To.Income.Ratio <- as.numeric(gsub("%","",loans$Debt.To.Income.Ratio))
#Factor State
loans$State <- as.factor(loans$State)
#Factor Home Ownership
loans$Home.Ownership <- as.factor(loans$Home.Ownership)
#Factor FICO Range
loans$FICO.Range <- as.factor(loans$FICO.Range)
#Factor Employment Lenght
loans$Employment.Length <- as.factor(loans$Employment.Length)
save(loans,dateDownloaded, file="../../data/loansData.rda")
loans640_724<-loans[loans$FICO.Range %in% levels(loans$FICO.Range)[1:17],]
```
### Look at the data set
```{r}
head(loans640_724)
summary(loans)
sapply(loans[1, ], class)
```

### Find out about missing values
```{r}
sum(is.na(loans))
```

Exploratory Analysis
---------------------------------------------------------------
### Make some univariate plots/summaries
```{r}
hist(loans640_724$Amount.Requested, 
     main="Histogram of Amount Requested",
     xlab="Amont Requested")
quantile(loans640_724$Amount.Requested)

hist(loans640_724$Amount.Funded.By.Investors,
     main="Histogram of Amount Funded By Investors",
     xlab="Amont Funded By Investors")
quantile(loans640_724$Amount.Funded.By.Investors)

hist(loans640_724$Interest.Rate,
     main="Histogram of Interest Rate",
     xlab="Interest Rate (%)")
quantile(loans640_724$Interest.Rate)

hist(loans640_724$Loan.Length,
     main="Histogram of Loan Lenght",
     xlab="Loan Lenght (months)")
quantile(loans640_724$Loan.Length)

hist(loans640_724$Debt.To.Income.Ratio,
     main="Histogram of Debt to Income Ratio",
     xlab="Debt to Income Ratio (%)")
quantile(loans640_724$Debt.To.Income.Ratio)

hist(loans640_724$Monthly.Income,
     main="Histogram of Monthly Income",
     xlab="Monthly Income ($)")
quantile(loans640_724$Monthly.Income, na.rm=TRUE)

hist(log(loans640_724$Monthly.Income),
     main="Histogram of log(Monthly Income)",
     xlab="Monthly Income (log($))")
quantile(log(loans640_724$Monthly.Income), na.rm=TRUE)

hist(loans640_724$Open.CREDIT.Lines,
     main="Histogram of Number of Credit Lines",
     xlab="Number of Credit Lines")
quantile(loans640_724$Open.CREDIT.Lines, na.rm=TRUE)

hist(loans640_724$Revolving.CREDIT.Balance,
     main="Histogram of Revolving Credit Balance",
     xlab="Revolving Credit Balance ($)")
quantile(loans640_724$Revolving.CREDIT.Balance, na.rm=TRUE )

hist(log(loans640_724$Revolving.CREDIT.Balance),
     main="Histogram of log(Revolving Credit Balance)",
     xlab="Revolving Credit Balance (log($)")
quantile(log(loans640_724$Revolving.CREDIT.Balance), na.rm=TRUE )

hist(loans640_724$Inquiries.in.the.Last.6.Months,
     main="Histogram of Number Inquiries in the last 6 months",
     xlab="Number of Inquiries in the last 6 months")
quantile(loans640_724$Inquiries.in.the.Last.6.Months, na.rm=TRUE)
```

### Make some univariate tables
```{r}
summary(loans640_724$Loan.Purpose)

summary(loans640_724$State)

summary(loans640_724$Home.Ownership)

summary(loans640_724$FICO.Range)

summary(loans640_724$Employment.Length)

```
### Plot Interest Rate vs Amount Requested (FICO Colored)
```{r fig.width=11, fig.height=8}
par(mar=c(4,4,2,0.5),col=palette(heat.colors(48)))
layout(matrix(c(1, 2, 3, 3), 2, 2, byrow=F), widths=c(3,1))
plot(loans640_724$Amount.Requested,loans640_724$Interest.Rate, pch=19,
     col=loans640_724$FICO.Range, 
     ylab="Interest Rate (%)", 
     xlab="Amount Request ($)", 
     main="Interest Rate vs Amount Requested (FICO Colored)",
     cex=0.7)
lmIRAR <- lm(loans640_724$Interest.Rate ~ loans640_724$Amount.Requested)
lines(loans640_724$Amount.Requested,lmIRAR$fitted,
      col="Black",lwd=3)
plot(loans640_724$Amount.Requested, lmIRAR$residuals, 
     col=loans640_724$FICO.Range, pch=19,
     ylab="lm No adjust Residuals", 
     xlab="Amount Request ($)", 
     main="Residuals Interest Rate ~ Amount Request (FICO Colored)",
     cex=0.7)
plot.new()
legend("center","groups",legend=sort(unique(loans640_724$FICO.Range)), 
       col=sort(unique(loans640_724$FICO.Range)),cex=0.7,pch=19, 
       title="FICO Score",ncol=1)
summary(lmIRAR)
```
### Plot Interest Rate vs Amount Funded By Investors (FICO Colored)
```{r fig.width=11, fig.height=6}
par(mar=c(4,4,2,0.5),col=palette(heat.colors(48)))
layout(matrix(c(1, 1, 2, 2), 2, 2, byrow=F), widths=c(3,1))
plot(loans640_724$Amount.Funded.By.Investors,loans640_724$Interest.Rate, pch=19, col=loans640_724$FICO.Range, ylab="Interest Rate (%)", xlab="Amount Funded By Investors ($)", main="Interest Rate vs Amount Funded By Investors (FICO Colored)")
lmIRAFBI <- lm(loans640_724$Interest.Rate ~ loans640_724$Amount.Funded.By.Investors)
lines(loans640_724$Amount.Funded.By.Investors,lmIRAFBI$fitted,
      col="black",lwd=3)
plot.new()
legend("center","groups",legend=sort(unique(loans640_724$FICO.Range)), 
       col=sort(unique(loans640_724$FICO.Range)),cex=0.7,pch=19, 
       title="FICO Score",ncol=1)
summary(lmIRAFBI)
```
### Plot Interest Rate vs Loan Lenght
```{r fig.width=11, fig.height=6}
boxplot(loans640_724$Interest.Rate ~ loans640_724$Loan.Length, ylab="Interest Rate(%)", xlab="Loan Lenght (months)", main="Plot Interest Rate vs Loan Lenght")
lmIRLL <- lm(loans640_724$Interest.Rate ~ loans640_724$Loan.Length)
summary(lmIRLL)
```
### Plot Interest Rate vs Loan Purpose
```{r fig.width=11, fig.height=6}
boxplot(loans640_724$Interest.Rate ~ loans640_724$Loan.Purpose, ylab="Interest Rate(%)", main="Plot Interest Rate vs Loan Porpose", las=3, cex.axis=0.6)
lmIRLP <- lm(loans640_724$Interest.Rate ~ loans640_724$Loan.Purpose)
summary(lmIRLP)
```
### Plot Interest Rate vs Debt to income ratio (FICO Colored)
```{r fig.width=11, fig.height=6}
par(mar=c(4,4,2,0.5),col=palette(heat.colors(48)))
layout(matrix(c(1, 1, 2, 2), 2, 2, byrow=F), widths=c(3,1))
plot(loans640_724$Debt.To.Income.Ratio,loans640_724$Interest.Rate, pch=19, col=loans640_724$FICO.Range, ylab="Interest Rate (%)", xlab="Debt to income ratio (%)", main="Interest Rate vs Debt to income ratio (FICO Colored)")
lmIRDIR <- lm(loans640_724$Interest.Rate ~ loans640_724$Debt.To.Income.Ratio)
lines(loans640_724$Debt.To.Income.Ratio, lmIRDIR$fitted,
      col="black",lwd=3)
plot.new()
legend("center","groups",legend=sort(unique(loans640_724$FICO.Range)), 
       col=sort(unique(loans640_724$FICO.Range)),cex=0.7,pch=19, 
       title="FICO Score",ncol=1)
summary(lmIRDIR)
````
### Plot Interest Rate vs State

```{r fig.width=11, fig.height=6}
boxplot(loans640_724$Interest.Rate ~ loans640_724$State, xlab="State", ylab="Interest Rate(%)", main="Plot Interest Rate vs State", las=3, cex.axis=0.6)
lmIRST <- lm(loans640_724$Interest.Rate ~ loans640_724$State)
summary(lmIRST)
````
### Plot Interest Rate vs Home Ownership
```{r fig.width=11, fig.height=6}
boxplot(loans640_724$Interest.Rate ~ loans640_724$Home.Ownership, xlab="Home Ownership", ylab="Interest Rate(%)", main="Plot Interest Rate vs Home Ownership", las=3, cex.axis=0.6)
lmIRHO <- lm(loans640_724$Interest.Rate ~ loans640_724$Home.Ownership)
summary(lmIRHO)
````
### Plot Interest Rate vs log(Monthly Income) (FICO Colored)
```{r fig.width=11, fig.height=6}
par(mar=c(4,4,2,0.5),col=palette(heat.colors(48)))
layout(matrix(c(1, 1, 2, 2), 2, 2, byrow=F), widths=c(3,1))
plot(log(loans640_724$Monthly.Income),loans640_724$Interest.Rate, pch=19, col=loans640_724$FICO.Range, ylab="Interest Rate (%)", xlab="log(Monthly Income ($))", main="Interest Rate vs log(Monthly Income) (FICO Colored)")
lmIRMI <- lm(loans640_724$Interest.Rate ~ log(loans640_724$Monthly.Income))
lines(log(loans[!is.na(loans640_724$Monthly.Income),"Monthly.Income"]), lmIRMI$fitted,
      col="black",lwd=3)
plot.new()
legend("center","groups",legend=sort(unique(loans640_724$FICO.Range)), 
       col=sort(unique(loans640_724$FICO.Range)),cex=0.7,pch=19, 
       title="FICO Score",ncol=1)
summary(lmIRMI)
````
### Plot Interest Rate vs FICO Range
```{r fig.width=11, fig.height=6}
boxplot(loans640_724$Interest.Rate ~ loans640_724$FICO.Range, xlab="FICO Range", ylab="Interest Rate(%)", main="Plot Interest Rate vs FICO Range", las=3, cex.axis=0.6)
lmIRFR <- lm(loans640_724$Interest.Rate ~ loans640_724$FICO.Range)
summary(lmIRFR)
````
### Plot Interest Rate vs Open Credit Lines
```{r fig.width=11, fig.height=6}
boxplot(loans640_724$Interest.Rate ~ loans640_724$Open.CREDIT.Lines, xlab="Open Credit Lines", ylab="Interest Rate(%)", main="Plot Interest Rate vs Open Credit Lines", las=3, cex.axis=0.6)
lmIROC <- lm(loans640_724$Interest.Rate ~ loans640_724$Open.CREDIT.Lines)
summary(lmIROC)
````
### Plot Interest Rate vs log(Revolving Credit Balance) (FICO Colored)
```{r fig.width=11, fig.height=6}
par(mar=c(4,4,2,0.5),col=palette(heat.colors(48)))
layout(matrix(c(1, 1, 2, 2), 2, 2, byrow=F), widths=c(3,1))
plot(log(loans640_724$Revolving.CREDIT.Balance+1),loans640_724$Interest.Rate, pch=19, col=loans640_724$FICO.Range, ylab="Interest Rate (%)", xlab="log(Revolving Credit Balance($))", main="Interest Rate vs log(Revolving Credit Balance) (FICO Colored)")
lmIRRCB <- lm(loans640_724$Interest.Rate ~ log(loans640_724$Revolving.CREDIT.Balance+1))
lines(log(loans[!is.na(loans640_724$Revolving.CREDIT.Balance),
            "Revolving.CREDIT.Balance"]+1), 
      lmIRRCB$fitted,
      col="black",lwd=3)
plot.new()
legend("center","groups",legend=sort(unique(loans640_724$FICO.Range)), 
       col=sort(unique(loans640_724$FICO.Range)),cex=0.7,pch=19, 
       title="FICO Score",ncol=1)
summary(lmIRRCB)
````
### Plot Interest Rate vs Inquiries in the last 6 months
```{r fig.width=11, fig.height=6}
boxplot(loans640_724$Interest.Rate ~ loans640_724$Inquiries.in.the.Last.6.Months, xlab="Number of Inquiries in the last 6 months", ylab="Interest Rate(%)", main="Plot Interest Rate vs Open Credit Lines", las=3, cex.axis=0.6)
lmIRIL6M <- lm(loans640_724$Interest.Rate ~ loans640_724$Inquiries.in.the.Last.6.Months)
summary(lmIRIL6M)
```
### Plot Interest Rate vs Employment Length
```{r fig.width=11, fig.height=6}
boxplot(loans640_724$Interest.Rate ~ loans640_724$Employment.Length, xlab="Employment Length", ylab="Interest Rate(%)", main="Plot Interest Rate vs Employment Length", las=3, cex.axis=0.6)
lmIREL <- lm(loans640_724$Interest.Rate ~ loans640_724$Employment.Length)
summary(lmIREL)
```

